Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach

### 2023-03-30 paper.tex
Merge sections 'Results' and 'Examples...'

In this contribution we continue to build upon foundational work of Ma and
Torquato \cite{ma2018SS} and recent findings of Samarin et al. \cite{our_SS} to
develop a robust and efficient approach to compute surface correlation functions
from digital 2D and 3D images. In partiuclar, here our aim is to establish a
direct verification of "discrete" methodology (that is easily parallelizable on
CPU and GPU architectures) using exact noiseless "continuous" approach for
arbitrary structures (still described by some continuous function). The rest of
the manuscript is organized as follows: in \cref{sec:def} we introduce some
major definitions of CFs necessary for establishing algorithms, in
\cref{sec:algo} we describe the all methodological details for "discrete" method
for digital images. Next, in \cref{sec:algo-precise} we explain in detail why
verification of computational approaches for digital images is currently
hampered, and then introduce novel exact "continuous" methodology in
\cref{sec:fss-2d}. The choice of verification cases is presented in
\textcolor{red}{Fix reference} followed by improvements in methodology of
Samarin et al. The paper concludes with discussion, outline and summary in
\cref{sec:summary}.

@@

\section{Examples of two-dimensional sets with smooth boundary}
\label{seq:examples}
In this section we will provide few examples of sets for which the developed
precise algorithm is applicable and compare the surface-surface function
calculated with the fast algorithm (\cref{sec:algo}) with results of the new
algorithm. We will use the filter $H$ (\ref{eq:filter-3x3}) for edge
detection along with a new filter $H'$ (\ref{eq:filter-7x7}) proposed in this
paper.

@@

\section{Results}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{images/ball-3d-ss.png}
  \caption[]{Surface-surface function calculated for a three-dimensional ball
    with radius $R = 0.24$ using filters $H$ and $H'$. Theoretical curve is
    provided for a comparison. Image dimensions are $450 \times 450 \times 450$.}
  \label{fig:fss-ball}
\end{figure}
With the new filter calculation of $F_{ss}$ becomes more precise, even for
three-dimensional images. This can be seen in comparison with theoretical values
for known cases like the one of a ball. On \cref{fig:fss-ball} there are plots
of $F_{ss}$ function calculated with filters $H$ and $H'$ along with a
theoretical curve. It can be verified that filter $H'$ is better indeed by
calculating
\begin{equation*}
  error = \sqrt{\frac{\sum_i (F_{ss}^{calc}(x_i) -
      F_{ss}^{theory}(x_i))^2}{\sum_i F_{ss}^{theory}(x_i)^2}}
\end{equation*}
where $x_i$ belong to a set
$(\epsilon, 2R - \epsilon) \cup (2R + \epsilon, 1)$ (here we assume image
dimensions to be $1 \times 1 \times 1$ length units). Taking
$\epsilon = 0.01$ we get $error = 0.006$ for filter $H$ and $error = 0.00017$
for filter $H'$.

>>>>

In this contribution we continue to build upon foundational work of Ma and
Torquato \cite{ma2018SS} and recent findings of Samarin et al. \cite{our_SS} to
develop a robust and efficient approach to compute surface correlation functions
from digital 2D and 3D images. In partiuclar, here our aim is to establish a
direct verification of "discrete" methodology (that is easily parallelizable on
CPU and GPU architectures) using exact noiseless "continuous" approach for
arbitrary structures (still described by some continuous function). The rest of
the manuscript is organized as follows: in \cref{sec:def} we introduce some
major definitions of CFs necessary for establishing algorithms, in
\cref{sec:algo} we describe the all methodological details for "discrete" method
for digital images. Next, in \cref{sec:algo-precise} we explain in detail why
verification of computational approaches for digital images is currently
hampered, and then introduce novel exact "continuous" methodology in
\cref{sec:fss-2d}. The choice of verification cases is presented in
\cref{sec:results} followed by improvements in methodology of Samarin et~al. The
paper concludes with discussion, outline and summary in \cref{sec:summary}.

@@

\section{Results}
\label{sec:results}
In this section we will provide few examples of sets for which the developed
precise algorithm is applicable and compare the surface-surface function
calculated with the fast algorithm (\cref{sec:algo}) with results of the new
algorithm. We will use the filter $H$ (\ref{eq:filter-3x3}) for edge
detection along with a new filter $H'$ (\ref{eq:filter-7x7}) proposed in this
paper.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{images/ball-3d-ss.png}
  \caption[]{Surface-surface function calculated for a three-dimensional ball
    with radius $R = 0.24$ using filters $H$ and $H'$. Theoretical curve is
    provided for a comparison. Image dimensions are $450 \times 450 \times 450$.}
  \label{fig:fss-ball}
\end{figure}
With the new filter calculation of $F_{ss}$ becomes more precise, even for
three-dimensional images. This can be seen in comparison with theoretical values
for known cases like the one of a ball. On \cref{fig:fss-ball} there are plots
of $F_{ss}$ function calculated with filters $H$ and $H'$ along with a
theoretical curve. It can be verified that filter $H'$ is better indeed by
calculating
\begin{equation*}
  error = \sqrt{\frac{\sum_i (F_{ss}^{calc}(x_i) -
      F_{ss}^{theory}(x_i))^2}{\sum_i F_{ss}^{theory}(x_i)^2}}
\end{equation*}
where $x_i$ belong to a set
$(\epsilon, 2R - \epsilon) \cup (2R + \epsilon, 1)$. Taking $\epsilon = 0.01$ we
get $error = 0.006$ for filter $H$ and $error = 0.00017$ for filter $H'$.
